\chapter{Criteria}
\label{chap:criteria}

Before a language analysis can take place, a set of criteria are needed. These criteria are the base from the evaluation, and are a mixture of several subjective criteria.

This chapter will present the criteria for the evaluation. Several criteria have been considered, where most are taken from known criteria for comparison \cite{design_criteria1} \cite{design_criteria2}, and others are found through discussion of the research questions.

\begin{description}[style=nextline]
\item[Readability] The language expresses itself through syntax as readability for the programmer. Readable code gives a greater understanding of the semantics as well of the nature of the code. For this project, readability is a vital criteria, as interest often follows understanding. Readability is hard to measure, as it is a subjective matter, mostly depending on the person writing code. The measurement of this quality will therefore be based on using code conventions and skilful coding.
\item[Writability] Writability is the ability to translate thoughts into code. It describes the expressivity of the code and the ease of writing, in a combination of quality and quantity. Writability is often found in the level of abstraction, and in the simplicity of the syntax.
\item[Observability] Observability is the level of feedback gained for a better understanding of the input. It is to what extend you can observe reactions to what you make. To measure the observability is to look into this level of feedback, and the informative value it provides.
\item[Trialability] The level of possibility for trial and error through coding is measured through trialability. This feature is measured by the level of feedback when an error occurs, how often feedback is given, and how easily a programmer can recover from a mistake.
\item[Learnability] As a language is learned, there are helping and hindering factors. These are measured through learnability, which is done by measuring the cost of learning the language and its environment, as well as the value gained through this cost.
\item[Reusability] The level of possibility for reusing code, through abstraction. This is measured both in quantity and quality, depending on the abstraction level and layer depth.
\item[Pedagogic Value] A programming language in itself can be easy to learn, but if it doesn't help the programmer in learning the basic concepts of programming, then there is no pedagogic value. This means the programming language should support the general programming concepts that are typical for common languages and coding conventions.
\item[Environment] The development environment plays an important part for novices, and it should provide help for the programmer in a simple, clear and manageable way. This criteria is based on the design and evaluation of interactive systems by David Benyon\cite[p. 225-250]{design_book}., and the evaluation will therefore be based on this.
\item[Documentation] The amount of documentation, as well as the informative value of this, is important for a novice, as a help for solving problems they cannot solve themselves.
\item[Uniformity] The consistency of appearance and behavior of language constructs. If the code does not look like any conventional language, the programmer will not learn the general approach to programming, and will have trouble moving on from this language. Constructs should be similar to commonly known syntaxes in the given paradigm, or preferably written in the exact same way.
\item[Miscellaneous] There is a possibility that other points of interest will be discovered during the comparison. These will be described in this category.
\end{description}

\section{Criteria Evaluation}
As the criteria are hard to measure in an objective way, the evaluation will be conducted through a subjective discussion. Each of the environments are meant for different purposes, as they are based on different paradigms. Taking this into account, we will create solutions to problems that are fit for each of the environments, and then compare these solutions in how all the environments handle and solve this problem, based on the criteria.