\chapter{Conclusion}
\label{chap:conclusion}

Through this report, we have analysed the field of novice programming and the problems regarding learning programming. We have investigated several languages for novices, and discussed their merits and shortcomings. There is a huge difference between visual and text based programming, which is addressed, where it was found that visual programming has an advantage in its ease of access and getting acquainted with, where the user for example do not use the keyboard for much. Text based programming, on the other hand, requires a better understanding of how a computer is operated, which could lead to a problem in the younger group of novices. We have analysed the field of teaching, and concluded that the teaching approach varies from country to country. The U.K. seems to have a good grasp on novice programming, where Denmark has some trouble, as the teaching is done by possibly unqualified teachers. The US have not made programming mandatory yet, but as an optional course, counting towards final grades. Furthermore, research is being made in how to incorporate programming in other courses.

We chose to continue the analysis with three novice languages and environments, being Scratch, BlueJ and DrRacket, for a language comparison. This was carried out as a subjective evaluation by the authors, with the intention of understanding the novice approach to programming. We analysed each language by a given set of criteria, which were criteria for evaluating programming languages in general, and for evaluating the interface design. Some sample pieces of code were made for comparison, which were made to match each of the paradigms representing the languages. After discussing each criteria for each language, with examples through the sample code, we compared the languages in each criteria. We concluded that a great difference lies in the targeting age. Scratch can work as a great introduction to programming down to the age of a primal school pupil. BlueJ serves as a great introduction to general purpose programming, but requires more effort and concentration, as well as external recourses for understanding Java. DrRacket can be hard, due to the different syntax. Furthermore, as with BlueJ, it needs external resources to function as a novice environment.

\input{./text/conclusion/discussion}

\input{./text/conclusion/further_works}